19Q1

### 2019-04-26 18Q4.Rmd
updated 18Q4 and 19Q1 Rmds, test CRISPR merging function and normalization testing

```{r, eval=FALSE}
cn_melt <- readRDS("./data_munging/rds/cn_melt_18Q4.rds")
ge_melt_18Q3 <- readRDS("./../htscreens_giant_files/ge_melt_18Q3.rds")
# ge_melt_18Q4 <- readRDS("./data_munging/rds/ge_melt_18Q4.rds")

@@

```{r}
maf_df <- readRDS("./data_munging/rds/maf_df_18Q4.rds")
test <- filter(maf_df, Hugo_Symbol %in% g2p_genes) %>% count(Hugo_Symbol, CCLE_Name)
test <- test[order(test$n, decreasing = TRUE),]
```

>>>>

```{r, eval=FALSE}
cn_melt <- readRDS("./data_munging/18Q4/cn_melt_18Q4.rds")
ge_melt_18Q3 <- readRDS("./../htscreens_giant_files/ge_melt_18Q3.rds")
# ge_melt_18Q4 <- readRDS("./data_munging/rds/ge_melt_18Q4.rds")

@@

```{r}
maf_df <- readRDS("./data_munging/18Q4/maf_df_18Q4.rds")
test <- filter(maf_df, Hugo_Symbol %in% g2p_genes) %>% count(Hugo_Symbol, CCLE_Name)
test <- test[order(test$n, decreasing = TRUE),]
```

### 2019-04-26 19Q1.Rmd
updated 18Q4 and 19Q1 Rmds, test CRISPR merging function and normalization testing

MergeOmicsG2P <- function(database, outfile) {
  # Merge mutation status, CN, and GE omics data with drugscreen results and metadata
  filt <- auc %>% filter(Database == database) %>% select(Cell_line, Histology, TCGA_classification, CCLE_Name, Broad_ID, Drug, AUC_IC50)
  print(paste0("Filtered for ", database, "."))
  colnames(filt)[colnames(filt) == "AUC_IC50"] <- "Score"

  # Get all G2Pgene-cell line combinations
  grid <- expand.grid("Cell_line" = unique(filt$Cell_line), "Gene" = g2p_genes)
  data <-  merge(filt, grid, by = "Cell_line", all.y = TRUE)
  print("Merged grid.")

  # Merge metadata
  data <-  merge(data, select(ccl_meta, Cell_line, Histology, CCLE_Name, Broad_ID, Primary.Disease), by = c("Cell_line", "Histology", "CCLE_Name", "Broad_ID"), all.x = TRUE)
  ## Add columns for G2P drug-gene and drug-gene-lineage sorting
  data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
  data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$Primary.Disease)
  print("Merged metadata.")

  # Merge mutation calls
  data <- merge(data, maf_df_filt, by = c("Gene", "Broad_ID"), all.x = TRUE)
  data$Mutation_Status <- ifelse(is.na(data$Mutation_Status), 0, data$Mutation_Status)
  ## Identify point mutations
  data$PointMutation <- ifelse(is.na(data$PointMutation), 1, data$PointMutation)
  ## Add color columns for plotting
  if(database == "CCLE") {
    data$Color <- ifelse(data$Mutation_Status == 0, "palegreen3", "springgreen4")
    data$Color <- factor(data$Color)
  }
  if(database == "CTRP") {
    data$Color <- ifelse(data$Mutation_Status == 0, "slategray3", "steelblue4")
    data$Color <- factor(data$Color)
  }
  if(database == "GDSC") {
    data$Color <- ifelse(data$Mutation_Status == 0, "paleturquoise3", "turquoise4")
    data$Color <- factor(data$Color)
  }
  print("Merged mutation calls.")

@@

`r nrow(filter(test, n != 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.
```{r, eval=FALSE}
# Load in Pozdeyev et al. (2016) data
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

@@

```{r, eval=FALSE}
crispr <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr) <- gsub(" .*", "", colnames(crispr))
colnames(crispr)[1] <- "Broad_ID"

# Melt CRISPR dataset for merging
crispr_melt <- melt(crispr, id.vars = "Broad_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Gene", value.name = "Score")

@@

@@

```{r}
test <- gdsc_ptmuts %>% count(Drug_Gene_Lin, Mutation_Status)
```

>>>>

MergeOmicsG2P <- function(database, outfile) {
  # Merge mutation status, CN, and GE omics data with drugscreen results and metadata
  if(database %in% c("CCLE", "CTRP", "GDSC")) {
    filt <- auc %>% filter(Database == database) %>% select(Cell_line, Histology, TCGA_classification, CCLE_Name, Broad_ID, Drug, AUC_IC50)
    print(paste0("Processing ", database, " data."))
    colnames(filt)[colnames(filt) == "AUC_IC50"] <- "Score"

    # Get all G2Pgene-cell line combinations
    grid <- expand.grid("Cell_line" = unique(filt$Cell_line), "Gene" = g2p_genes)
    data <-  merge(filt, grid, by = "Cell_line", all.y = TRUE)
    print("Merged grid.")
    # Merge metadata
    data <-  merge(data, select(ccl_meta, Cell_line, Histology, CCLE_Name, Broad_ID, Primary.Disease), by = c("Cell_line", "Histology", "CCLE_Name", "Broad_ID"), all.x = TRUE)
    ## Add columns for G2P drug-gene and drug-gene-lineage sorting
    data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
    data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$Primary.Disease)
    print("Merged metadata.")
  if(database == "CRISPR") {
    print(paste0("Processing ", database, " data."))
    # Melt CRISPR dataset for merging
    data <- melt(crispr, id.vars = "Broad_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Gene", value.name = "Score")
    # Merge cell line metadata
    data <- merge(data, select(ccl_meta, Cell_line, Histology, CCLE_Name, Broad_ID, Primary.Disease), by = "Broad_ID", all.x = TRUE)
    data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
    data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$Primary.Disease)
    print("Merged metadata.")
  }

  # Merge mutation calls
  data <- merge(data, maf_df_filt, by = c("Gene", "Broad_ID"), all.x = TRUE)
  data$Mutation_Status <- ifelse(is.na(data$Mutation_Status), 0, data$Mutation_Status)
  ## Identify point mutations
  data$PointMutation <- ifelse(is.na(data$PointMutation), 1, data$PointMutation)
  # Summarize number of mutant and Wildtype cell lines
  data_muts_summ <- data %>% group_by(Gene) %>% summarize(N_Wildtype = length(Mutation_Status) - sum(Mutation_Status == 1), N_Mutant = sum(Mutation_Status == 1))
  ## Merge test results back into full dataset, which restores information lost in the summarization
  data <- merge(data_muts_summ, data, by = "Gene")
  ## Add color columns for plotting
  if(database == "CCLE") {
    data$Color <- ifelse(data$Mutation_Status == 0, "palegreen3", "springgreen4")
    data$Color <- factor(data$Color)
  }
  if(database == "CTRP") {
    data$Color <- ifelse(data$Mutation_Status == 0, "slategray3", "steelblue4")
    data$Color <- factor(data$Color)
  }
  if(database == "GDSC") {
    data$Color <- ifelse(data$Mutation_Status == 0, "paleturquoise3", "turquoise4")
    data$Color <- factor(data$Color)
  }
  if(database == "CRISPR") {
    data$Color <- ifelse(data$Mutation_Status == 0, "thistle", "darkorchid")
    data$Color <- factor(data$Color)
  }
  print("Merged mutation calls.")

@@

`r nrow(filter(test, n != 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.

**NEED TO RUN CTRP OVER WEEKEND**
```{r, eval=FALSE}
# Load in Pozdeyev et al. (2016) data
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

@@

```{r, eval=FALSE}
crispr_all <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr_all) <- gsub(" .*", "", colnames(crispr_all))
colnames(crispr_all)[1] <- "Broad_ID"

# Melt CRISPR dataset for merging
crispr_melt <- melt(crispr_all, id.vars = "Broad_ID", measure.vars = colnames(crispr_all)[2:ncol(crispr_all)], variable.name = "Gene", value.name = "Score")

@@

Test merging and filtering functions on CRISPR data:
```{r, eval=FALSE}
crispr <- MergeOmicsG2P(database = "CRISPR", outfile = "./../htscreens_giant_files/crispr_data_19Q1_g2pgenes.rds")
crispr_ptmuts <- FilterForPointMuts(dataframe = crispr, outfile = "./../htscreens_giant_files/crispr_data_19Q1_g2pgenes_ptmuts.rds")
```

@@

# SANDBOX

Filter for BRCA1/2 mutations to see if we catch non-germline mutations/CN variants:
```{r}
test <- filter(maf_df, Gene == "BRCA1" | Gene == "BRCA2")
test2 <- filter(cn_melt, Gene == "BRCA1" | Gene == "BRCA2")
```

Test out standardization/normalization methods on Pozdeyev data. This yields evidence that the AUC values are already normalized.
```{r, eval=FALSE}
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ccle_all$AUC_IC50_standardized <- scale(ccle_all$AUC_IC50, center = TRUE, scale = TRUE)
ccle_all$AUC_IC50_normalized <- scale(ccle_all$AUC_IC50, center = min(ccle_all$AUC_IC50[!is.na(ccle_all$AUC_IC50)]), scale = max(ccle_all$AUC_IC50[!is.na(ccle_all$AUC_IC50)]) - min(ccle_all$AUC_IC50[!is.na(ccle_all$AUC_IC50)]))
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
ctrp_all$AUC_IC50_standardized <- scale(ctrp_all$AUC_IC50, center = TRUE, scale = TRUE)
ctrp_all$AUC_IC50_normalized <- scale(ctrp_all$AUC_IC50, center = min(ctrp_all$AUC_IC50[!is.na(ctrp_all$AUC_IC50)]), scale = max(ctrp_all$AUC_IC50[!is.na(ctrp_all$AUC_IC50)]) - min(ctrp_all$AUC_IC50[!is.na(ctrp_all$AUC_IC50)]))
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")
gdsc_all$AUC_IC50_standardized <- scale(gdsc_all$AUC_IC50, center = TRUE, scale = TRUE)
gdsc_all$AUC_IC50_normalized <- scale(gdsc_all$AUC_IC50, center = min(gdsc_all$AUC_IC50[!is.na(gdsc_all$AUC_IC50)]), scale = max(gdsc_all$AUC_IC50[!is.na(gdsc_all$AUC_IC50)]) - min(gdsc_all$AUC_IC50[!is.na(gdsc_all$AUC_IC50)]))

auc_test <- bind_rows(ccle_all, ctrp_all, gdsc_all)

ggplot(data = auc_test) +
  geom_histogram(aes(x = AUC_IC50)) +
  facet_wrap(~ Database, ncol = 3, scales = "free_y")
ggplot(data = auc_test) +
  geom_histogram(aes(x = AUC_IC50_normalized)) +
  facet_wrap(~ Database, ncol = 3, scales = "free_y")
ggplot(data = auc_test) +
  geom_histogram(aes(x = AUC_IC50_standardized)) +
  facet_wrap(~ Database, ncol = 3, scales = "free_y")
```
